### 2021-11-15 input/proposal.tex
starting with xas, small changes here and there
about purge

\subsection{Scoped broadcast \TODO{Is it so important?}}
\label{subsec:scoped}

@@

\begin{theorem}[\label{theo:bef}\underline{E}ventual \underline{F}orwarding
    of \underline{B}est $(EFB) \implies CP$] Assuming reliable
  communication links where a correct \process $q$ eventually receives
  the message $m$ \underline{s}ent to it by a \process $p$ ($s_{pq}(m)
  \implies r_{q}(m)$), \processes eventually reach consistent
  partitioning if each \process eventually \underline{f}orwards
  ($f_p(m) \implies \forall \langle p, q\rangle \in E: s_{pq}(m)$) its
  best known partition.
\end{theorem}

@@

Unfortunately, stale control information as stated in
Definition~\ref{def:lwo} may impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issues with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending ($s_{pq}(m) \rightarrow s_{pq}(m') \implies r_q(m) \rightarrow
r_q(m')$). Both $a$ and $d$ add then delete their partition
concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, $b$ receives,
delivers, and forwards $\alpha_a^2$. In Figure~\ref{fig:problemC},
both $c$ and $d$ deliver the message about Partition $P_a$, for they
did not have any known partition at receipt time. On the contrary, $b$
delivers $\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks $a$'s removal notification
$\delta_a$. It never reaches $c$ nor $d$. Also, $c$ does not deliver
$\alpha_d$ and $\delta_d$ since it already delivered it. The system
converges to an inconsistent state where some \processes assume they
belong to a partition while it does not exist anymore. The system
needs to purge stale $\alpha$ messages that stop $\delta$ messages
from reaching nodes.
%The system needs to
% purge its stale messages.

%% Naive propagation of $\alpha$ and $\delta$ messages
%% is insufficient to guarantee consistent partitioning when any \process
%% can add or delete its partition at any time. \Processes need to
%% exploit their local knowledge gathered over message receipts to allow
%% up-to-date messages to propagate to relevant \processes.

\begin{theorem}[\label{theo:dcp}EFB$^+$: EFB $\wedge$ (Purge $\rightarrow$ EFB) $\implies$
    \underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
%
When a \process can \texttt{Add} and \texttt{Del}ete its partition, to
ensure consistent partitioning, eventual forwarding of best (see
Theorem~\ref{theo:bef}) additionally requires
\begin{inparaenum}[(i)]
\item eventual purging of stale messages ($S(m)
  \implies (d_q(m) \implies \eventually d_q(m) \rightarrow
  d_q(m'))$);
\item then the eventual forwarding of best such that $m'$ ends up
  originating from the best up-to-date source for $q$.
\end{inparaenum}
\end{theorem}

\begin{proof}
  %% To reach consistent partitioning, the last message delivered by
  %% every process is not stale and comes from its respective best
  %% source: $d(m)$
  Assuming EFB as default propagation behavior where \processes
  deliver and forward the best sum of weights they received. We must
  prove that despite stale messages, the last delivery of every
  \process is the message that followed the shortest path to its
  closest source.
  \begin{asparadesc}
  \item [EFB without Purge:] From Figure~\ref{fig:problemD}, a
    \process may receive messages in such an order ($d_b(\alpha_d^1)
    \rightarrow r_b(\delta_a) \rightarrow d_b(\delta_d)$) that it does
    not forward messages required by other \processes
    ($\delta_a$). The last delivery of some nodes ($c$ and $d$) is
    wrong ($\alpha_a^3$ instead of $\delta_a$).
    % \Processes never
    % reach consistent partitioning.
    %
    %% Figure~\ref{fig:proofA} depicts the generic case where a \process
    %% $b$ precludes itself and subsequent \processes such as $c$ of
    %% delivering the message from their respective best source
    %% $a$. \Process $b$ received and delivered a message from a
    %% partition that has since then been deleted but with a better sum
    %% of weights than newly received ones ($\alpha_d^y <
    %% \alpha_a^{x'}$). It forever discards such a message
    %% $\alpha_a^{x'}$.
  \item [Purge not followed by EFB:]
    %% also transiting messages
    Assuming that every \process removes control information as soon
    as it becomes stale ($S(\alpha_a) \implies d_b(\alpha_a^x)
    \rightarrow d_b(\delta_a)$) and cannot deliver stale messages
    anymore ($S(\alpha_a^x) \implies d_c(m) \rightarrow \ldots
    \rightarrow d_c(m' \not\in \alpha_a)$), staleness does not impair
    up-to-date messages propagation. However, even in such settings,
    depending on receipt order ($d_b(\alpha_a^{x}) \rightarrow
    r_b(\alpha_c^{y : y > x}) \rightarrow d_b(\delta_b)$), a \process
    may never receive again, hence deliver and forward a message that
    was received and discarded before ($\alpha_c^{y}$). When such a
    message contains its best up-to-date partition, a \process may
    never end up in its best partition.
    %    Therefore, \processes do not reach
    % consistent partitioning.
  \item [Purge then EFB:] Assuming the eventual and definitive removal
    of stale messages at each \process ($S(\alpha_a) \implies
    (d_b(\alpha_a) \implies \eventually d_b(\alpha_a) \rightarrow
    d_b(m') \not\rightarrow d_b(\alpha_a))$); and assuming that every
    removal eventually triggers the forwarding of best delivered
    messages ($d_c(\delta_x) \implies \eventually f_d(\alpha_y)$),
    \processes may deliver other stale messages that will be 
    eventually removed to never be delivered again. In the worst case,
    all \processes deliver and remove all stale messages to ultimately
    deliver and forward only up-to-date messages ($\eventually d_c(m')
    \implies \neg S(m')$). As in Theorem~\ref{theo:bef}, they
    eventually deliver and forward their best up-to-date message.

@@

\begin{figure}
  \centering\input{input/figproofB.tex}
  \caption{\label{fig:proofB}Stale $\alpha$ messages may stop $\delta$
    messages from reaching \processes with the corresponding
    partition. Consistent partitioning requires that all \processes
    propagate $\delta$ messages as often as possible (I); detect
    possible inconsistencies when parents' $\alpha$ messages
    contradict history or state (II); notify \processes of possible
    inconsistencies (III).}
\end{figure}

>>>>

\subsection{Scoped broadcast}
\label{subsec:scoped}

@@

\begin{theorem}[\label{theo:efb}\underline{E}ventual \underline{F}orwarding
    of \underline{B}est $(EFB) \implies CP$] Assuming reliable
  communication links where a correct \process $q$ eventually receives
  the message $m$ \underline{s}ent to it by a \process $p$ ($s_{pq}(m)
  \implies r_{q}(m)$), \processes eventually reach consistent
  partitioning if each \process eventually \underline{f}orwards
  ($f_p(m) \implies \forall \langle p, q\rangle \in E: s_{pq}(m)$) its
  best known partition.
\end{theorem}

@@

Unfortunately, stale control information as stated in
Definition~\ref{def:lwo} may impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
\TODO{Rework here.}
Figure~\ref{fig:problem} highlights such consistency issues with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending ($s_{pq}(m) \rightarrow s_{pq}(m') \implies r_q(m) \rightarrow
r_q(m')$). Both $a$ and $d$ add then delete their partition
concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, $b$ receives,
delivers, and forwards $\alpha_a^2$. In Figure~\ref{fig:problemC},
both $c$ and $d$ deliver the message about Partition $P_a$, for they
did not have any known partition at receipt time. On the contrary, $b$
delivers $\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks $a$'s removal notification
$\delta_a$. It never reaches $c$ nor $d$. Also, $c$ does not deliver
$\alpha_d$ and $\delta_d$ since it already delivered it. The system
converges to an inconsistent state where some \processes assume they
belong to a partition while it does not exist anymore. The system
needs to purge stale $\alpha$ messages that stop $\delta$ messages
from reaching nodes.

\begin{definition}[\label{def:purge}Message purge]
  A system purges a message $P(m)$ when every \node $q$ that delivered
  $m$ eventually delivers another message $m'$ never followed by the
  delivery of the former message $m$: $d_q(m) \implies \eventually
  (d_q(m) \rightarrow d_q(m'))$.
\end{definition}

\begin{theorem}[\label{theo:dcp}EFB$^+$: EFB $\wedge$ (Purge $\rightarrow$ EFB) $\implies$
    \underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
%
  When a \process can \texttt{Add} and \texttt{Del}ete its partition,
  to ensure consistent partitioning, eventual forwarding of best
  (Theorem~\ref{theo:efb}) additionally requires
\begin{inparaenum}[(i)]
\item eventual purging of stale messages ($S(m) \implies P(m)$), then
\item the eventual forwarding of best (Definition~\ref{theo:efb}).
\end{inparaenum}
\end{theorem}

\begin{proof}
  %% To reach consistent partitioning, the last message delivered by
  %% every process is not stale and comes from its respective best
  %% source: $d(m)$
  Assuming EFB as default propagation behavior where \processes
  deliver and forward the best sum of weights they received. We must
  prove that despite stale messages, the last delivery of every
  \process is the message that followed the shortest path to its
  closest source.
  \begin{asparadesc}
  \item [EFB without Purge:] From Figure~\ref{fig:problemD}, a
    \process may receive messages in such an order ($d_b(\alpha_d^1)
    \rightarrow r_b(\delta_a) \rightarrow d_b(\delta_d)$) that it does
    not forward messages required by other \processes
    ($\delta_a$). The last delivery of some nodes ($c$ and $d$) is
    wrong ($\alpha_a^3$ instead of $\delta_a$).
    % \Processes never
    % reach consistent partitioning.
    %
    %% Figure~\ref{fig:proofA} depicts the generic case where a \process
    %% $b$ precludes itself and subsequent \processes such as $c$ of
    %% delivering the message from their respective best source
    %% $a$. \Process $b$ received and delivered a message from a
    %% partition that has since then been deleted but with a better sum
    %% of weights than newly received ones ($\alpha_d^y <
    %% \alpha_a^{x'}$). It forever discards such a message
    %% $\alpha_a^{x'}$.
  \item [Purge not followed by EFB:]
    %% also transiting messages
    Assuming that every \process removes control information as soon
    as it becomes stale ($S(\alpha_a) \implies d_b(\alpha_a^x)
    \rightarrow d_b(\delta_a)$) and cannot deliver stale messages
    anymore ($S(\alpha_a^x) \implies d_c(m) \rightarrow \ldots
    \rightarrow d_c(m' \not\in \alpha_a)$), staleness does not impair
    up-to-date messages propagation. However, even in such settings,
    depending on receipt order ($d_b(\alpha_a^{x}) \rightarrow
    r_b(\alpha_c^{y : y > x}) \rightarrow d_b(\delta_b)$), a \process
    may never receive again, hence deliver and forward a message that
    was received and discarded before ($\alpha_c^{y}$). When such a
    message contains its best up-to-date partition, a \process may
    never end up in its best partition.
    %    Therefore, \processes do not reach
    % consistent partitioning.
  \item [Purge then EFB:] Assuming the eventual and definitive removal
    of stale messages at each \process ($S(\alpha_a) \implies
    (d_b(\alpha_a) \implies \eventually d_b(\alpha_a) \rightarrow
    d_b(m') \not\rightarrow d_b(\alpha_a))$); and assuming that every
    removal eventually triggers the forwarding of best delivered
    messages ($d_c(\delta_x) \implies \eventually f_d(\alpha_y)$),
    \processes may deliver other stale messages that will be 
    eventually removed to never be delivered again. In the worst case,
    all \processes deliver and remove all stale messages to ultimately
    deliver and forward only up-to-date messages ($\eventually d_c(m')
    \implies \neg S(m')$). As in Theorem~\ref{theo:efb}, they
    eventually deliver and forward their best up-to-date message.

@@

%% \begin{figure}
%%   \centering\input{input/figproofB.tex}
%%   \caption{\label{fig:proofB}Stale $\alpha$ messages may stop $\delta$
%%     messages from reaching \processes with the corresponding
%%     partition. Consistent partitioning requires that all \processes
%%     propagate $\delta$ messages as often as possible (I); detect
%%     possible inconsistencies when parents' $\alpha$ messages
%%     contradict history or state (II); notify \processes of possible
%%     inconsistencies (III).}
%% \end{figure}
